---
title: "ECON 444 Problem Set 2"
subtitle: "Josepeh Sepich Feb 14"
output:
  pdf_document:
    toc: false
    toc_depth: 2
    number_sections: true
    keep_tex: false
---

```{r include=FALSE}
library(tidyverse)
```

# Problem 1

Problem Constraints

* Market Demand: $Q(p) = 1000 * p^x$
* Marginal Cost (Supply): $p = 2$
* $x = -3$

What price would a monopolist choose?

In order to maximize profits a monopolist would set marginal revenue equal to marginal cost:

\[\text{MR} = \text{MC}\]

We already have that marginal cost is a constant, but what is marginal revenue? We can
determine this value by defining total revenue:

\[\text{TR} = p * Q\]

We know that marginal revenue is the first derivative of total revenue, since it is 
the additional revenue for each additional unit.

\[\text{MR} = \frac{d\text{TR}}{dQ}\]
\[\frac{d\text{TR}}{dQ} = \frac{d(p * Q)}{dQ}\]

Here we must use our market demand equation and we can put $p$ in terms of $Q$:

\[Q = 1000 * p^x\]
\[\frac{Q}{1000} = p^x\]
\[p = (\frac{Q}{1000})^{1/x}\]

Plugging this back in we can continue.

\[\frac{d((\frac{Q}{1000})^{1/x} * Q)}{dQ}\]

According to our problem constraints $x = 3$.

\[\frac{d((\frac{Q}{1000})^{-1/3} * Q)}{dQ} = \frac{d(\frac{Q^{2/3}}{1000^{-1/3}})}{dQ} = \frac{2}{3}(\frac{Q}{1000})^{-1/3}\]
\[\text{MR} = \frac{2}{3}(\frac{Q}{1000})^{-1/3}\]

Now solve for monopolist price:

\[\text{MR} = \text{MC}\]
\[\frac{2}{3}(\frac{Q}{1000})^{-1/3} = 2\]
\[\frac{Q}{1000} = 3^{-3}\]
\[Q = 1000(3)^{-3} \approx 37.04\]

To get the price of sale, plug this quantity into demand (translated into terms of price above):

\[p = (\frac{Q}{1000})^{-1/3} = (\frac{37.04}{1000})^{-1/3} = 3\]

The monopolist would choose a **price of 3**. Now let's calculate profit if $\text{FC} = 25$.

\[\pi = \text{TR} - \text{TC} = p * Q - (FC + Q * \text{MC})\]
\[\pi = 3 * 37.04 - (25 + 37.04 * 2) = 12.03704\]

The monopoly's profit level is **$12.04**.

We can reinforce this with the following plot:

```{r}
Q <- seq(0, 50, 0.01)
demand <- (Q / 1000) ^ (-1 / 3)
mc <- rep(2, length(Q))
mr <- (2 / 3) * (Q / 1000) ^ (-1 / 3)

data <- data.frame(Q = Q, demand = demand, mc = mc, mr = mr)

data %>%
  ggplot(aes(x = Q)) +
  geom_line(aes(y = demand)) +
  geom_line(aes(y = mc)) +
  geom_line(aes(y = mr)) +
  ylim(0, 10)
```

# Problem 2

## Part a

## Part b

## Part c

# Problem 3

## Part a

## Part b

## Part c

# Problem 4

## Part a

## Part b



