---
title: "ECON 444 Problem Set 5"
subtitle: "Joseph Sepich April 19"
output:
  pdf_document:
    toc: false
    toc_depth: 2
    number_sections: true
    keep_tex: false
---

```{r include=FALSE}
library(tidyverse)
```

# Problem 1

Problem Constrains:

* Cournot Competition (quantity static)
* Market Demand: $P = 10 - 2Q$
* $\text{MC}_1 = 2$
* $\text{MC}_2 = 4$
* $Q = q_1 + q_2$

## Part a

To find best response behavior of each firm, we acknowledge that the firms wish to set MR equal to MC to optimize profits.

\[MR = p*q\]
\[MR_1 = \frac{d((10 - 2(q_1 + q_2)) * q_1)}{dq_1} = \frac{d(10q_1 - 2q_1^2-2q_1q_2)}{dq_1} = 10 - 4q_1 - 2q_2\]
\[MR_2 = \frac{d((10 - 2(q_1 + q_2)) * q_2)}{dq_2} = \frac{d(10q_2 - 2q_2^2-2q_2q_1)}{dq_2} = 10 - 4q_2 - 2q_1\]

Best Response for firm 1:

\[MR_1 = MC_1\]
\[10 - 4q_1 - 2q_2 = 2\]
\[4q_1 = 10 - 2q_2 - 2\]
\[q_1^* = \frac14(8 - 2q_2)\]

Best Response for firm 2:

\[MR_2 = MC_2\]
\[10 - 4q_2 - 2q_1 = 4\]
\[4q_2 = 10 - 2q_1 - 4\]
\[q_2^* = \frac14(6 - 2q_1)\]

## Part b

To find equilibrium we must solve our system of best response equations.

\[4q_1 = 8 - 2 (\frac14(6 - 2q_1))\]
\[16q_1 = 32 - 2(6 - q_1)\]
\[16q_1 = 32 - 12 + 2q_1\]
\[14q_1 = 20\]
\[q_1^* = \frac{20}{14} = \frac{10}7\]

We can plug this back in to solve for firm 2.

\[q_2 = \frac14(6 - 2\frac{20}{14})\]
\[56q_2 = 84 - 40\]
\[56q_2 = 44\]
\[q_2* = \frac{11}{14}\]

\[Q^* = q_1^* + q_2^* = \frac{20 + 11}{14} = \frac{31}{14}\]
\[P^* = 10 - 2Q^* = 10 - \frac{31}{7} = \frac{39}{7} \approx 5.57\]

## Part c

\[\text{Markup:}\frac{p-c}{p}\]

\[m_1 = \frac{p-MC_1}{p} = \frac{5.57 - 2}{5.57} \approx 0.64\]
\[m_1 = \frac{p-MC_2}{p} = \frac{5.57 - 4}{5.57} \approx 0.28\]

The markup in equilibrium for firm 1 is **0.64** and **0.28** for firm 2.

## Part d

\[P = 10-2Q = 10-2(q_1 + q_2)\]

Firm 1 residual demand according to best response:

\[P = 10-2(q_1 + q_2*) = 10 - 2q_1 - 2(\frac{11}{14}) = 10 - \frac{11}{7} - 2q_1\]
\[q_1 = \frac12(8.43 - P)\]

Firm 2 residual demand according to best response:

\[P = 10-2(q_2 + \frac{20}{14}) = 10 - \frac{20}{7}-2q_2\]
\[q_2 = \frac12(7.14 - P)\]

The slope of residual demand is **-2** for both firm 1 and 2. (Inverse market demand curve)

## Part e

Recall the elasticity of demand.

\[\epsilon = \frac{dQ}{dP}\frac{P}{Q}\]

\[\epsilon = \frac{-P}{2Q} = \frac{-5.57}{2(q)}\]

\[\epsilon_1 = \frac{-5.57}{2(1.43)} = -1.95\]
\[\epsilon_2 = \frac{-5.57}{2(0.786)} = -3.54\]

The residual demand elasticity for firm 1 is **-1.95** and **-3.54** for firm 2. The firm with the smaller market share has larger residual demand elasticity. If the firm supplying less decides to supply more, then there will be a greater effect on decrease in prices.

## Part f

The markup does not equal the inverse of residual demand elasticity for either firm.

# Problem 2

Using the equations from derivation we can create the sample demand curves:

\[BR_1(q_2) = \frac{a - c_1 - bq_2}{2b} = q_1^m - \frac12q_2\]
\[BR_2(q_1) = \frac{a - c_2 - bq_1}{2b} = q_2^m - \frac12q_1\]

The $q_1$ intercept for firm 1's BR curve would be $q_1^m$ and for firm two would be $2q_2^m$. The $q_2$ intercept for each curve would be $2q_1^m$ and $q_2^m$ respectively. Firm 2 will never produce if their best response curve lies under the best response curve for firm 1. This means that the monopoly quantity amount of firm 1 must be more than double the equilibrium quntity of firm 2. An example can be seen below with a = 10, b = 1, c1 = 2, c2 = 8.

```{r echo=FALSE}
a <- 10
b <- 1
c1 <- 2
c2 <- 8

q1m <- (a - c1) / (2*b)
q2m <- (a - c2) / (2*b)

q1 <- seq(0, 5, 0.1)
BR1 <- 2*q1m - 2*q1
BR2 <- q2m - q1 / 2

data <- data.frame(q1 = q1, br = BR1, label = rep("BR1", length(q1)))
data <- data %>%
  rbind(data.frame(q1 = q1, br = BR2, label = rep("BR2", length(q1))))

data %>%
  ggplot(aes(x = q1, y = br, color = label)) +
  geom_line() + ylim(0,8)
```

# Problem 3

## Part a

## Part b

# Problem 4

## Part a

## Part b

## Part c

# Problem 5

## Part a

## Part b









